---
title: "nam218grads"
output: github_document
---

This package simplifies access to the [NCEP NOMADS grads server](http://nomads.ncep.noaa.gov:80/dods/nam) for the [NAM](https://www.ncdc.noaa.gov/data-access/model-data/model-datasets/north-american-mesoscale-forecast-system-nam) forecast models.

## Requirements

 + [R v4+](https://www.r-project.org/)
 + [rlang](https://CRAN.R-project.org/package=rlang)
 + [dplyr](https://CRAN.R-project.org/package=dplyr)
 + [sf](https://CRAN.R-project.org/package=sf)
 + [stars](https://CRAN.R-project.org/package=stars)
 + [ncdf4](https://CRAN.R-project.org/package=ncdf4)

## Installation

```
remotes::install_github("BigelowLab/nam218grads")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Usage

### Retrieve a URL for a resource

Retrieve a URL for a resource by date, like yesterday's forecast at 0000H. Note that the
NCEP/NOMADS grads server only maintains the current forecast plus the prior 6.

```{r get_uri, message = FALSE}
library(stars)
library(nam218grads)
uri <- grads_uri(date = Sys.Date()-1, product = "nam_00z")
uri
```

### Get a handle to a resource

It's just a [ncdf4](https://CRAN.R-project.org/package=ncdf4) object, but we wrapped common
tasks in friendly functions.

```{r grads_open}
x <- open_grads(uri)
# File http://nomads.ncep.noaa.gov:80/dods/nam/nam20210303/nam_00z (NC_FORMAT_CLASSIC):
# 
#      141 variables (excluding dimension variables):
#         float absvprs[lon,lat,lev,time]   
#             _FillValue: 9.99900026055401e+20
#             missing_value: 9.99900026055401e+20
#             long_name: ** (1000 850 700 500 250) absolute vorticity [1/s] 
#         float no4lftx180_0mb[lon,lat,time]   
#             _FillValue: 9.99900026055401e+20
#             missing_value: 9.99900026055401e+20
#             long_name: ** 180-0 mb above ground best (4 layer) lifted index [k] 
#      ...
#    4 global attributes:
#        title: NAM every 3 hours fcst staring from 00Z03mar2021, download Mar 03 02:38 UTC
#        Conventions: COARDS GrADS
#        dataType: Grid
#        history: Wed Mar 03 22:27:39 GMT 2021 : imported by GrADS Data Server 2.0
```

You can explore the contents of the resource.

```{r times}
times <- get_time(x)
times
```

```{r levels}
levels <- get_lev(x)
levels
```

```{r variables}
vars <- get_varnames(x)
vars
```

```{r dimnames}
dims <- get_dims(x)
dims
```


### Retrieve data

#### As an array
Retrieving data will return contiguous arrays of data. We provide a function `get_var_array` that expects the slab coordinates in integer run-length encoding for each dimension as `[start_index, count]`. This function returns an array of cell values - for range of pressure levels at different times.

```{r get_array}
a <- nam218grads::get_var_array(x, "tmpprs", 
                                list(lon = c(110, 100),
                                     lat = c(150, 75),
                                     lev = c(3, 5),
                                     time = c(1, 12)))
str(a)
```


#### As a stars object
The `get_var` function is a convenient wrapper around the `get_var_array` function.  Users specify the slab limits with 'real world' units, and a [`stars`](https://r-spatial.github.io/stars/) object is returned (although returning an array is an option).

```{r get_stars}
s <- nam218grads::get_var(x, "tmpprs",
                          bb = c(-77.0, -51.5, 37.9, 56.7),
                          time = times[1:12],
                          lev = levels[3:5])
s
```

You can also retrieve multiple variables. Note in the example below that `ugrd10m` and `vgrd10m` (wind component speeds at 10m above surface) do not have a `lev` dimension so we can ignore the `lev` argument.

```{r multiple_vars}
winds <- nam218grads::get_var(x, c("ugrd10m", "vgrd10m"),
                          bb = c(-77.0, -51.5, 37.9, 56.7),
                          time = times[5:7])
winds
```

#### Close the handle

The resource is a ncdf4 object - be sure to close it when done.

```{r close_it}
close_grads(x)
```

### Plot the levels at different times

Let's use the temperature at various pressure levels in variable `s` for some plotting. First time slice...

```{r plot_time_1}
if (require(maps)){
  time_string <- format(times[1], "%Y-%m-%d %H:%M:%S")
  overlay_fun <- function() {
    map(col = 'red', add = TRUE)
    mtext(time_string, side = 1, line = -1)
  }
  plot(s, axes = TRUE, hook = overlay_fun)
}
```

And the last time slice...

```{r plot_time_12}
if (require(maps)){
  time_string <- format(times[12], "%Y-%m-%d %H:%M:%S")
  plot(s[,,,,12], axes = TRUE, hook = overlay_fun)
}
```
